---
title: "Introduction to Code Blocks"
description: "This is my reading note about features of code blocks in Quarto."
---

## Code Execution Options

Code execution options can be used when you want to, e.g., regulate the code output, hide the code piece, and so on. Read [the official guide about execution options](https://quarto.org/docs/computations/execution-options.html) for more detail.

## Interactive Code Annotations

Interactive code annotation is a combination of inline comment and bullet points. The display of code annotation can be controled by the variable `code-annotations`, which contains the following options: 

* below (default)
* hover
* select
* none

Below is an example of using code annotation from [the official guide about code annotation](https://quarto.org/docs/authoring/code-annotation.html).

```{r load packages}
#| echo: false
#| output: false

# (install &) load packages
pacman::p_load(
    conflicted,
    tidyverse
)

# handle function conflicts
conflicts_prefer(dplyr::filter)
conflicts_prefer(dplyr::lag)
```

```{r load data with annotation}
pacman::p_load(palmerpenguins)
penguins |> # <1>
    mutate( # <2>
        bill_ratio = bill_depth_mm / bill_length_mm, # <2>
        bill_area  = bill_depth_mm * bill_length_mm # <2>
    ) # <2>
```

1. Take `penguins`, and then,
2. add new columns for the bill ratio and bill area.
