---
title: "Gaining a First Impression on Your Data"
description: "This article covers some methods of inspecting the data when you get it at the first time."
---

## Basic Methods

In this example we use the dataset `PlantGrowth`. We can simply gain a first impression on the data by simply calling `PlantGrowth` directly.

```{r direct call on the dataset}
PlantGrowth
```

We begin with the conventional approach: loading the data and save it as a `data.frame`.

```{r load as df}
df <- PlantGrowth
```

We can print some basic information about this `data.frame` using the `str()` method.

```{r str representation of df}
str(df)
```

Using `summary()` is a very better way of gaining a comprehensive view on the data.

```{r summary of df}
summary(df)
```

We can also get an overview on values of every columns.

```{r show the weight column}
df$weight
```

```{r show the group column}
df$group
```

In addition, we can also show the first or the last rows.

```{r head and tail}
head(df)
tail(df)
```

## Inspect the Data using the `{dplyr}` Package

[`dplyr`](https://dplyr.tidyverse.org//) is one of the most popular package from the `tidyverse` universe. Before we proceed, we need to load this package.

```{r load dplyr}
pacman::p_load(dplyr)
```

### `dplyr::glimpse()`

The `dplyr::glimpse()` method provides a similar output as the `str()` method.

```{r using glimpse()}
glimpse(PlantGrowth)
```

### Using `tibble`

A `tibble` is a more efficient version of a `data.frame`. A `tibble` is based on `data.frame` and in most cases its behavior still mimics a `data.frame`. Despite, here are some benefits of using `tibble` over `data.frame`:

1. It tells the data type of each column.
2. Instead of printing all the data, it only prints a limited numbers of rows and columns, so that the console won't be flooded by too many text.
3. Missing values and negative values are printed in red.

By using the function `dplyr::as_tibble()`, we can convert a `df` into a `tibble`.

```{r load a df as a tibble()}
tbl <- as_tibble(df)
tbl
```

### Print a `tibble`

By default 10 rows are printed.

```{r default behavior of print()}
print(tbl) # To 10 rows
```

To print more rows we can:
```{r print 20 rows}
print(tbl, n = 20) # To 20 rows
```

To print all rows we can:
```{r print all rows}
print(tbl, n = Inf) # To print all rows
```

### Transform a `tibble` into a `data.frame`

A `tibble` can be transformed into a `data.frame`.

```{r transform a tibble into a data.frame}
tbl %>% as.data.frame()
```

## Compare `data.frame` and `tibble`

Here is a comparison of calling some common functions on a `tibble` and on a `data.frame`.

:::: {.columns}
::: {.column width="49%"}
```{r}
class(tbl)
str(tbl)
head(tbl)
summary(tbl)
tbl$weight
tbl$group
```
:::

::: {.column width="2%"}
<!-- empty column to create gap -->
:::

::: {.column width="49%"}
```{r}
class(df)
str(df)
head(df)
summary(df)
df$weight
df$group
```
:::
::::  
<!-- end of columns -->

## What's next?

Reading the numerical data is intuitive, despite, we prefer "seeing" the data. Navigate to [the introduction to `ggplot2`](ggplot2-basic.html) for more information about plotting the data.

In addition, you can also navigate to [this article](playing-with-data.html) to gain knowledge about how you can use your data and extract useful information.
