---
title: "Package Loading and Conflicting Management"
description: "This page contains the information about packages loading and conflict resolving"
---

## Installing and Loading Packages

### Brief Overview of Differnt Package Loading Methods

When using built-in R functions

```{r using built-in package loading functions}
#| eval: false

install.packages(packagename) # install package every time, even though the package is already installed
library(packagename) # If the package is not found, an error will be raised
```

When using `pacman::p_load()`, it only installs packages when the package is not found on the system. Note here we don't load the package `{pacman}`. Instead, we only use the `p_load()` function directly from this package.

```{r using pacman}
#| eval: false

pacman::p_load("packagename1", "packagename2") 
```

### Installing Packages from a user-specified CRAN Mirror Site

When you install packages using `install.packages()` method, you may encounter the following message:

```default
--- Please select a CRAN mirror for use in this session ---
```

You can either follow the GUI to choose your desire mirror site (which requires TCL/TK), or you can  simply pass an argument `repos` without the GUI tool, e.g.:

```{r select mirror site during function call}
#| eval: false

install.packages('RMySQL', repos='http://cran.us.r-project.org')
```

Here is the full list of [CRAN mirror sites](https://cran.r-project.org/mirrors.html).

::: {.callout-note}
Read more information from [this Stack Overflow discussion](https://stackoverflow.com/q/11488174).
:::

## Conflict Resolving

By default, R uses the function of the package that was loaded **THE LAST**. Instead of relying on this built-in mechanism, it's a good practice to manage package confilicts explicitly.

E.g. there are two common functions with the same name but come from different packages: 

* `dplyr::filter()`
* `stats::filter()`

We can use the `{conflicted}` package to deal with the conflict.

```{r conflict management}
# (install &) load packages
pacman::p_load(
    conflicted, 
    tidyverse)

# handle function conflicts
conflicts_prefer(dplyr::filter) 
```

::: {.callout-note title="External resources"}
* https://schmidtpaul.github.io/dsfair_quarto/ch/misc/usefulthings.html#conflicted
:::

## Removing Packages

[Here's a good example](https://stackoverflow.com/a/26633042) of removing a package along with its dependencies.
