---
title: "Playing with Your Data using `{dplyr}`"
description: "This article covers some common functions in the `{dplyr}` package for processing the data."
---

## Load the Data

In the following example, we use a partial of the `PlantGrowth` dataset.

```{r as tibble}
pacman::p_load(dplyr)
tbl <- as_tibble(PlantGrowth)
tbl2 <- head(tbl)
```

::: {.callout-note}
Read [this article](first-impression-data.html) for more information about the `PlantGrowth` dataset itself.
:::

## `mutate()`

baseR

```{r}
tbl2$new <- tbl2$weight + 2
```

Create new column

```{r}
tbl2 <- tbl2 %>% mutate(new = weight + 2)
```

Replace the value in-place

```{r}
tbl2 <- tbl2 %>% mutate(weight = weight + 2)
```

Note that when we export it to excel, the backtick won't shown in the excel file.

```{r}
tbl2 %>%
    mutate(
    `Name with space` = "Hello!", 
    number10 = 10
    )
```

For the 3

```{r}
tbl2 %>%
  mutate(v1 = 1, v2 = 2, v3 = 3) %>%
  mutate(
    across(starts_with("v"), ~ .x + 20))
```

## `select()`

```{r}
tbl3 <- tbl2 %>% 
    mutate(var1 = 1, var2 = 2, var3 = "text", var4 = "word")
tbl3
```

```{r}
tbl3 %>% 
    select(group, var1, var4)
```

```{r}
tbl3 %>% 
    select(group, var1:var4)
```

```{r}
tbl3 %>% 
    select(-group)
```

```{r}
tbl3 %>% 
    select(where(is.numeric))
```

Select *literally* everything but move the column `var` to the front.
```{r}
tbl3 %>% 
    select(var1, everything())
```

## `filter()`

```{r}
tbl %>% 
  filter(weight > 6 & group == "trt2")
```

```{r}
tbl %>% 
  filter(weight > 6 | weight < 4)
```

Now we want to keep only `trt1` and `trt2` observations

Option 1:

```{r}
tbl %>% 
  filter(group == "trt1" | group == "trt2")
```

Option 2:

```{r}
tbl %>% 
  filter(group %in% c("trt1", "trt2"))
```

Option 3:

```{r}
tbl %>% 
  filter(group != "ctrl")
```

```{r}
tbl %>% 
  filter(weight > mean(weight))
```

## `arrange()`

Firstly we gather the data. Note that we make use of `slice()`, whose function is to select data of discrete ranges.

```{r}
tbl4 <- tbl %>%
  slice(1:3, 11:13, 21:23) 
# this keeps only rows 1,2,3,11,12,13,21,22,23
```

Sort by weight (default: ascending)

```{r}
tbl4 %>%
  arrange(weight) 
```

Sort by weight, but descending:

```{r}
tbl4 %>%
  arrange(desc(weight))
```

Firstly sort by "group". Within each "group" we sort by "weight".  

```{r}
tbl4 %>%
  arrange(group, weight)
```

Custom order (e.g. sort not alphabetically):

```{r}
myorder <- c("trt1", "ctrl", "trt2")

tbl4 %>% 
  arrange(
    match(group, myorder), 
    weight
  )
```

## `summarise()`

```{r}
tbl %>% 
  # no group_by 
  summarise(my_mean = mean(weight))
```
But why not use the built-in `mean()` function? (i.e. in this case `mean(tbl$weight)`).

But `summary()` shines when combining with other functions. E.g. we can get the mean of each control groups:
```{r}
tbl %>% 
  group_by(group) %>% 
  summarise(my_mean = mean(weight))
```

We can also elaborate by combining with multiple functions:

```{r}
tbl %>% 
  group_by(group) %>% 
  summarise(
    Mean = mean(weight),
    StdDev = sd(weight),
    Min = min(weight),
    Median = median(weight),
    Max = max(weight),
    n_Obs = n(),
  )
```

We can even group by more than one groups

```{r}
tbl %>% 
  mutate(larger5 = case_when(
    weight > 5 ~ "yes",
    weight < 5 ~ "no"
  )) %>% 
  group_by(group, larger5) %>% 
  summarise(
    n_Obs = n(),
    Mean = mean(weight)
  )
```

## What's next?

Here are [some tests](tidyverse-exercise.html) for you.
