---
title: "Read & Export Data"
description: "This covers topics regarding loading and exporting data from external source."
---

## Read a csv file using the `tidyverse` packages

`readr::read_csv()` is a part of the `tidyverse` universe.

```{r load readr}
pacman::p_load(readr)
```

You can use this function to read a csv file not only from your local storage, but also from an online source.

```{r read csv from an online source}
path <- "https://raw.githubusercontent.com/yuchen-xue/Learn-R-Quarto/main/content/data/Clewer&Scarisbrick2001.csv"
dat <- read_csv(path) # use path from above
```

### Difference between `read.csv()` and `read_csv()`
Its output of `readr::read_csv()` is a `tibble`, whereas the native `read.csv()` function produces a `data.frame` object.

```{r}
#| eval: false

read.csv(path_to_csv) # base R import as data.frame
read_csv(path_to_csv) # load as `tidyverse::tibble` 
```

### Suppressing the message regarding column types

When loading the csv file, there might be a message goes like:

```default
Specify the column types or set `show_col_types = FALSE` to quiet this message.
```

According to [this Stack Overflow discussion](https://stackoverflow.com/q/55687401), you have two options. Either -- as suggested by the message -- pass the `show_col_types = FALSE` argument, or add information about the separator.

:::: {.columns}
::: {.column width="49%"}

```{r}
dat <- read_csv(
    path, show_col_types = FALSE)
```

:::

::: {.column width="2%"}
<!-- empty column to create gap -->
:::

::: {.column width="49%"}

```{r}
dat <- read_csv(
    path, col_types = cols())
```

:::

::::  
<!-- end of columns -->

## Read a Excel `.xlsx` Files

### Method 1: using `readxl::read_excel()`
The conventional `readxl::read_excel()` function can be used for reading  a `.xlsx` file from local storage. But it fails to read a file directly from an online source. Thus downloading the data is necessary before reading a `.xlsx` file using `readxl::read_excel()` function.

```{r using readxl::read_excel()}
pacman::p_load(readxl, httr)
xlsx_path <- "https://raw.githubusercontent.com/yuchen-xue/Learn-R-Quarto/main/content/data/vision_fixed.xlsx"

# Download the `.xlsx` file as a tempfile and loaded it locally
GET(xlsx_path, write_disk(tf <- tempfile(fileext = ".xlsx")))
xlsx_dat <- read_excel(tf, 2L)

# Remove the tempfile
unlink(tf)
xlsx_dat
```

### Method 2: using `read.xlsx()` from package `openxlsx`

`read.xlsx()` from package `openxlsx` can read a `.xlsx` file not only locally, but also directly from the online source. Thus it's the recommended method of file loading. 

```{r using openxlsx}
pacman::p_load(openxlsx)
xlsx_path <- "https://raw.githubusercontent.com/yuchen-xue/Learn-R-Quarto/main/content/data/vision_fixed.xlsx"
xlsx_dat <- read.xlsx(xlsx_path)
xlsx_dat
```

::: {.callout-note}
Learn more about loading a `.xlsx` file from an online resource by following [this Stack Overflow discussion](https://stackoverflow.com/q/41368628).
:::

## Write a csv file using the `tidyverse` packages

Learn more from this [doc](https://readr.tidyverse.org/reference/write_delim.html).
