---
title: "Read & Export Data"
description: "This is my reading note about loading and exporting data."
---

## Read a csv file using the `tidyverse` packages

`readr::read_csv()` is a part of the `tidyverse` universe. Its output is a `tibble`, instead of a `data.frame`.

::: {.callout-important}
Don't be confused with the native `read.csv()` function, which produces a `data.frame` object.
:::

```{r}
pacman::p_load(readr)
path <- "../data/Clewer&Scarisbrick2001.csv"
dat <- read_csv(path) # use path from above
```

### Suppressing the output

You will notice from the output there's a message:

```default
Specify the column types or set `show_col_types = FALSE` to quiet this message.
```

### Suppressing using function arguments

According to [this Stack Overflow discussion](https://stackoverflow.com/q/55687401), you have two options. Either -- as suggested by the message -- pass the `show_col_types = FALSE` argument, or add information about the separator.

:::: {.columns}
::: {.column width="49%"}

```{r}
dat <- read_csv(
    path, show_col_types = FALSE)
```

:::

::: {.column width="2%"}
<!-- empty column to create gap -->
:::

::: {.column width="49%"}

```{r}
dat <- read_csv(
    path, col_types = cols())
```

:::

::::  
<!-- end of columns -->

## Read a Excel `.xlsx` Files

```{r}
pacman::p_load(readxl)
xlsx_path <- "../data/vision_fixed.xlsx"
xlsx_dat <- read_excel(xlsx_path)
xlsx_dat
```

## Write a csv file using the `tidyverse` packages

Learn more from this [doc](https://readr.tidyverse.org/reference/write_delim.html).
