---
title: "Exercise of `{tidyverse}`"
code-fold: true
---

## Introduction

This page contains several tests. The output of each answer is shown in order to help you understand the questions. The code itself is hidden and you can unfold them when needed.

```{r load packages & data}
#| output: false
#| echo: false

# (install &) load packages
pacman::p_load(
    broom,
    conflicted,
    modelbased,
    tidyverse
)

# handle function conflicts
conflicts_prefer(dplyr::filter)
conflicts_prefer(dplyr::select)
```

## Exercise 1

Given the data:

```{r load_tibble}
#| code-fold: false

tbl <- as_tibble(PlantGrowth)
tbl
```

Use the original `tbl`, create a new column "Kg" and assume that the weight columns is in tons that you convert to Kg.
Afterwards, delete the weight column.
Keep only values smaller than 5000 Kg.
Sort the data so that the largest Kg are on top

```{r}
tbl %>%
    mutate(Kg = weight * 1000) %>%
    select(-weight) %>%
    filter(Kg < 5000) %>%
    arrange(desc(Kg))
```

## Exercise 2
Given the data `dat`:

```{r}
#| code-fold: false
#| output: false

path <- "https://raw.githubusercontent.com/SchmidtPaul/dsfair_quarto/master/data/Mead1993.csv"
dat <- read_csv(path, col_types = cols()) # use path from above
```

```{r}
#| code-fold: false

dat
```

Find out the average, minimum and maximum yields per variety using `summarise()`. Arrage as the highest value on the top.

```{r}
dat %>%
    group_by(variety) %>%
    summarise(
        avg_yield = mean(yield),
        min_yield = min(yield),
        max_yield = max(yield)
    ) %>%
    arrange(desc(avg_yield))
```
