---
title: "One Way CRD Analysis"
description: "this is a reading note of [this tutorial](https://schmidtpaul.github.io/dsfair_quarto/ch/exan/simple/crd_mead1993.html)."
---

## Load Data & Packages
```{r load packages & data}
#| output: false

# (install &) load packages
pacman::p_load(
  broom,
  conflicted,
  modelbased,
  tidyverse)

# handle function conflicts
conflicts_prefer(dplyr::filter) 
conflicts_prefer(dplyr::select)

# load data and change variable into the correct data type
path <- "https://raw.githubusercontent.com/SchmidtPaul/dsfair_quarto/master/data/Mead1993.csv"
dat <- read_csv(path) %>% 
  mutate(variety = as.factor(variety))
```

## Inspect the data

```{r inspect data}
dat %>% 
  group_by(variety) %>% 
  dlookr::describe(yield) %>% 
  select(2:sd, p00, p100) %>%
  arrange(desc(mean))
```

In the data, only "yield" is true numerial. "variety" is a factor and has 4 levels.

Below is the plot of the data.

```{r plot data}
ggplot(data = dat) +
  aes(y = yield, x = variety) +
  geom_point() + 
  # ggtitle("Hello Test") +  # this a short cut function
  labs(
    title = "my title", 
    subtitle = "mysubtitile", 
    caption = "and caption"
  ) + 
  # theme(
  #   axis.line.x = element_line(color = "green", linetype = "dotted")
  # )
  theme_classic() + 
  scale_y_continuous(
    name = "Yield [Kg/ha]", 
    limits = c(0, NA), # y-axis starts at 0 and ends at ggplot-default, i.e. the highest point.
    
    # fixed value is possible
    # expand = c(0, 0), 

    expand = expansion(mult = c(0, 0.1)), # but defined dynamically is better
  )
```

## Modelling

```{r lm modelling}
mod <- lm(yield ~ variety, data = dat)
mod
```

Since "variety" is a factor, instead of one single slope, we've got one estimate for the effect brought by each level. Here's the interpretation of the coefficient of our lm model:

* Intercept -- the baseline, which shows the yield of planting solely melon of variety 1 is $20.49$ kg/ha.
* The rest of the value -- the individual deviation for each further variety

A summary of prediction based on the model:

* pred of var 1: $20.49 + 0 = 20.49$ kg/ha
* pred of var 2: $20.49 + 16.9133 = 37.40$ kg/ha
* pred of var 3: $20.49 - 0.9983 = 19.49$ kg/ha
* pred of var 4: $20.49 + 9.4067 = 29.89$ kg/ha

## AVOVA (F-test)

Null hypothesis `H_0`: no matter if var 1, var 2, var 3 or var 4 is chosen, the yield stays the same.
Alternative hypothesis `H_A`: at least one of them makes difference.

```{r ANOVA}
anova(mod)
```

Since $p < 0.001$, we believe it's extreme unlikely to find a sample that support the null hypothesis. Thus we believe in with the alternative hypothesis instead, i.e. our lm model has statistic significance.
